Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-01-15 TEBenchmarkPaper/article.tex
More table justification

% For interestingness table
\newcommand{\iE}{\textbf{E}}
\newcommand{\iN}{\textbf{N}}
\newcommand{\iS}{\textbf{S}}
\newcommand{\iA}{\textbf{A}}
\newcommand{\iC}{\textbf{C}}
\newcommand{\iU}{\textbf{U}}
\newcommand{\tIFF}{if-and-only-if}
\newcommand{\tNE}{non-exists}
\newcommand{\tIMP}{implies}
\newcommand{\tEQ}{equations}
\newcommand{\tRow}[1]{#1 \\ \hline}

@@

{\bf Applicability} depends on the number of models in which a
conjecture holds. The system of Bagai et al conjectures the
non-existence of objects, and hence favours statements with zero
applicability. Other systems treat ap- plicability as a positive
aspect: the more applicable the statement, the more interesting it is.

@@

\begin{table}
  \centering
  \begin{tabular}{ |l|l|c|c|c|c|c|c| }
    \hline
    \multirow{2}{*}{\textbf{Program}}                     &
    \multirow{2}{*}{\textbf{Conjecture Types}}            &
    \multicolumn{6}{c}{\textbf{Interestingness Measures}} \\ \hhline{~~------}
    \tRow{            &                    & \iE & \iN & \iS & \iA & \iC & \iU}
    \tRow{AM          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{GT          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{Graffiti    & inequalities       &   X &   X &   X &     &   X &   X}
    \tRow{Bagai et al & \tNE               &     &   X &     &   X &   X &    }
    \tRow{HR          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{QuickSpec   & \tEQ               &   X &   X &     &     &   X &    }
    \tRow{Speculate   & conditional \tEQ   &   X &   X &     &   A &   X &    }
    \tRow{IsaCoSy     & \tEQ               &   X &   X &     &     &   X &   X}
    \tRow{IsaScheme   & \tEQ               &   X &   X &     &     &   X &   X}
  \end{tabular}
  \caption{Classification of ATE systems from \cite{colton2000notion}, extended
    to those compared in \cite{claessen2013automating} (QuickSpec is the
    conjecture generation component of HipSpec). The interestingness measures
    are \iE{}mpirical plausibility, \iN{}ovelty, \iS{}urprisingness,
    \iA{}pplicability, \iC{}omprehensibility (low complexity) and \iU{}tility.}
  \label{table:colton}
\end{table}

@@

We summarise the criteria in Table 1 and state which were used in key
ATF/ATE systems. The entries for AM, GT, Graffiti, Bagai et al. and HR are based
on this survey (the latter is Colton's own system). We extend this analysis to
some more recent systems in the remaining rows (based on our understanding of
their function): QuickSpec, Speculate, IsaCoSy and IsaScheme.

All four of these systems use counterexample checking and a small-to-large
search order, favouring empirical plausibile and comprehensibile output. IsaCoSy
ensures novelty using a constraint system, such that no generated expression can
be simplified using a previously-found equation. The other systems use post-hoc
filters based on unification and rewriting (based on the Knuth-Bendix completion
algorithm) to discard expressions or equations which are special-cases of some
other, more general result.

>>>>

% For interestingness table
\newcommand{\iE}{\textbf{E}}
\newcommand{\iN}{\textbf{N}}
\newcommand{\iS}{\textbf{S}}
\newcommand{\iA}{\textbf{A}}
\newcommand{\iC}{\textbf{C}}
\newcommand{\iU}{\textbf{U}}
\newcommand{\tIFF}{if-and-only-if}
\newcommand{\tNE}{non-exists}
\newcommand{\tIMP}{implies}
\newcommand{\tEQ}{equations}
\newcommand{\tINE}{inequalities}
\newcommand{\tCON}{conditional}
\newcommand{\tRow}[1]{#1 \\ \hline}

@@

{\bf Applicability} depends on the number of models in which a
conjecture holds. The system of Bagai et al conjectures the
non-existence of objects, and hence favours statements with zero
applicability. Other systems treat applicability as a positive
aspect: the more applicable the statement, the more interesting it is.

@@

\begin{table}
  \centering
  \begin{tabular}{ |l|l|c|c|c|c|c|c| }
    \hline
    \multirow{2}{*}{\textbf{Program}}                     &
    \multirow{2}{*}{\textbf{Conjecture Types}}            &
    \multicolumn{6}{c}{\textbf{Interestingness Measures}} \\ \hhline{~~------}
    \tRow{            &                    & \iE & \iN & \iS & \iA & \iC & \iU}
    \tRow{AM          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{GT          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{Graffiti    & \tINE              &   X &   X &   X &     &   X &   X}
    \tRow{Bagai et al & \tNE               &     &   X &     &   X &   X &    }
    \tRow{HR          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{QuickSpec   & \tEQ               &   X &   X &     &   X &   X &    }
    \tRow{Speculate   & \tCON \tEQ / \tINE &   X &   X &     &   X &   X &    }
    \tRow{IsaCoSy     & \tEQ               &   X &   X &     &   X &   X &   X}
    \tRow{IsaScheme   & \tEQ               &   X &   X &     &   X &   X &   X}
  \end{tabular}
  \caption{Classification of ATE systems from \cite{colton2000notion}, extended
    to those compared in \cite{claessen2013automating} (QuickSpec is the
    conjecture generation component of HipSpec). The interestingness measures
    are \iE{}mpirical plausibility, \iN{}ovelty, \iS{}urprisingness,
    \iA{}pplicability, \iC{}omprehensibility (low complexity) and \iU{}tility.}
  \label{table:colton}
\end{table}

@@

We summarise the criteria in Table 1 and state which were used in key
ATF/ATE systems. The entries for AM, GT, Graffiti, Bagai et al. and HR are based
on this survey (the latter is Colton's own system). We extend this analysis to
some more recent systems in the remaining rows (based on our understanding of
their function): QuickSpec, Speculate~\cite{braquehais2017speculate}, IsaCoSy
and IsaScheme.

All four of these systems use testing to check for counterexamples, which
ensures results are empirically plausible and applicable (conditions are
satisfiable, types are inhabited, etc.). They also use a small-to-large search
order, ensuring more easily-comprehensibile conjectures are explored first.

IsaCoSy ensures novelty using a constraint system, such that no generated
expression can be simplified using a previously-found equation. The other
systems use post-hoc filters based on unification and rewriting rules
(derived via the Knuth-Bendix completion algorithm) to discard expressions or
equations which are special-cases of some other, more general result.
